### 2021-11-09 manuscript/manuscript.Rmd

---
title: "Morality, politics, and cooperation"
output: bookdown::html_document2
date: '`r format(Sys.time(), "%d %B, %Y")`'
bibliography: references.bib
knit: worcs::cite_all
---

```{r setup, include=FALSE}
# We recommend that you prepare your raw data for analysis in 'prepare_data.R',
# and end that file with either open_data(yourdata), or closed_data(yourdata).
# Then, uncomment the line below to load the original or synthetic data
# (whichever is available), to allow anyone to reproduce your code:
library("worcs")
knitr::opts_chunk$set(echo = FALSE, results = "hide", message = FALSE, warning = FALSE)
options(knitr.kable.NA = '')
library(ggplot2)
library(tidySEM)
library(lavaan)
library(semTools)
library(kableExtra)
library(bookdown)
library(psych)
dat <- load_data(to_envir = FALSE)
source("../scales_list.R")
scales_list <- lapply(scales_list, function(i){
  tmp <- lapply(names(i), function(j){ paste0(j, "_", 1:length(i[[j]]))})
  names(tmp) <- names(i)
  tmp
  })
set.seed(1)
cfas <- lapply(names(scales_list), function(cnt){
  results <- t(sapply(scales_list[[cnt]], function(scal){
    out <- cfa(paste0("F =~ ", paste0(scal, collapse = "+")), std.lv = TRUE, data = dat[[cnt]])
    paran <- fa.parallel(dat[[cnt]][, scal])
    tab <- table_results(out, columns = NULL, digits=3)
    tab <- tab[-nrow(tab), ]
    loadns <- as.numeric(tab$est_std)[tab$op == "=~"]
    r2s <- lavInspect(out, "r2")
    omga <- semTools::reliability(out, what = "omega")[1,1]
    c(Subscale = gsub("_\\d$", "", scal[1]),
      Items = length(scal),
      n = out@Data@nobs[[1]],
      formatC(unlist(table_fit(out)[1, c("chisq", "cfi", "tli", "rmsea", "srmr")]), digits = 3, format = "f"), 
      formatC(c(min_load = min(loadns), max_load = max(loadns), min_r2 = min(r2s), max_r2 = max(r2s), omega = omga), digits = 3, format = "f"),
      Reliability = tidySEM:::interpret(omga),
      Factors = fa.parallel(dat[[cnt]][,scal]) $nfact
      )
  }))
})
names(cfas) <- names(scales_list)
```

## Necessary deviations from preregistration

When attempting to conduct the preregistered analyses,
we noticed that the model did not converge in one of the datasets (NL).
We examined individual CFAs for the included scales to determine potential
sources of misspecification.
These analyses, summarized below, indicate that several scales in that dataset
displayed evidence of not being unidimensional,
had poor reliability, or had no factors with Eigenvalues greater than would be expected by random chance.

```{r tabscale, results = "asis"}
invisible(lapply(names(cfas), function(cntr){
  print({kbl(cfas[[cntr]], caption = paste0("Scale reliability ", cntr)) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
})}))
```

To understand these issues better, we performed exploratory factor analysis on scales that showed indications of being non-unidimensional in at least one country.

```{r notunidim, results = "asis"}
allscales <- do.call(c, scales_list)
nfacs <- as.integer(do.call(c, lapply(cfas, function(i){i[, 'Factors']})))
check_these <- nfacs > 1
these_scales <- unique(do.call(c, lapply(cfas, function(i){i[, 'Subscale']}))[check_these])
check_these <- grep(paste0("^.{2}\\.(", paste0(these_scales, collapse = "|"), ")$"), names(allscales))
these_scales <- allscales[check_these]
nfacs <- nfacs[check_these]
nfacs[nfacs == 0] <- 2
out <- mapply(function(scal, nfac){
  fa(dat[[substring(scal, first = 1, last = 2)]][these_scales[[scal]]], nfac)
}, scal = names(these_scales), nfac = rep(2, length(these_scales)))

these_scales <- unique(do.call(c, lapply(cfas, function(i){i[, 'Subscale']}))[check_these])
for(scal in these_scales){
  print(kbl(do.call(cbind,
  lapply(grep(scal, names(out), value = TRUE), function(i){
    tmp <- data.frame(unclass(out[[i]][["loadings"]]))
    names(tmp) <- paste0(substring(i, 1,2), " Factor ", 1:ncol(tmp))
    rownames(tmp) <- paste0("Item ", 1:nrow(tmp))
    tmp
  })
), digits = 2, caption = paste0("Factor loadings for ", scal))|>
  kable_styling(bootstrap_options = c("striped", "hover")))
}
```

These factor analyses suggest that items 3-5 of sepa_soc loaded most consistently on the same factor.
Similarly, items 3-5 of sepa_eco loaded most consistently on the same factor.
For secs_soc, only items 4-5 loaded consistently high on the same factor.
For secs_eco, no items loaded consistently high on the same factor.
We removed items not consistently loading on one factor,
and dropped secs_eco entirely.

```{r, include=FALSE}
use_scales <- lapply(scales_list, function(i){
  thenames <- grepl("sepa_(soc|eco)", names(i))
  if(any(thenames)){
    i[which(thenames)] <- lapply(i[which(thenames)], `[`, 3:5)
  }
  i
})
use_scales <- lapply(use_scales, function(i){
  thenames <- grepl("secs_soc", names(i))
  if(any(thenames)){
    i[which(thenames)] <- lapply(i[which(thenames)], `[`, 4:5)
  }
  i
})
use_scales$us$secs_eco <- NULL
use_scales$nl$secs_eco <- NULL

set.seed(1)
cfas <- lapply(names(use_scales), function(cnt){
  results <- t(sapply(use_scales[[cnt]], function(scal){
    out <- cfa(paste0("F =~ ", paste0(scal, collapse = "+")), std.lv = TRUE, data = dat[[cnt]])
    paran <- fa.parallel(dat[[cnt]][, scal])
    tab <- table_results(out, columns = NULL, digits=3)
    tab <- tab[-nrow(tab), ]
    loadns <- as.numeric(tab$est_std)[tab$op == "=~"]
    r2s <- lavInspect(out, "r2")
    omga <- semTools::reliability(out, what = "omega")[1,1]
    c(Subscale = gsub("_\\d$", "", scal[1]),
      Items = length(scal),
      n = out@Data@nobs[[1]],
      formatC(unlist(table_fit(out)[1, c("chisq", "cfi", "tli", "rmsea", "srmr")]), digits = 3, format = "f"), 
      formatC(c(min_load = min(loadns), max_load = max(loadns), min_r2 = min(r2s), max_r2 = max(r2s), omega = omga), digits = 3, format = "f"),
      Reliability = tidySEM:::interpret(omga),
      Factors = invisible(fa.parallel(dat[[cnt]][,scal])$nfact)
      )
  }))
})
names(cfas) <- names(use_scales)

meas_inv <- sapply(use_scales$dk, function(scal){
  tmpdat <- rbind(tryCatch(cbind(dat$nl[, scal], Country = "NL"), error = function(e) NULL),
                  tryCatch(cbind(dat$dk[, scal], Country = "DK"), error = function(e) NULL),
                  tryCatch(cbind(dat$us[, scal], Country = "US"), error = function(e) NULL))
  conf <- cfa(paste0("F =~ ", paste0(scal, collapse = "+")), std.lv = TRUE, data = tmpdat, group = "Country")
  metr <- cfa(paste0("F =~ ", paste0(scal, collapse = "+")), std.lv = TRUE, data = tmpdat, group = "Country", group.equal = "loadings")
  tst <- anova(conf, metr)
  unlist(tst[2, c("Chisq diff", "Df diff", "Pr(>Chisq)")])
})
```

```{r tabscaleuse, results = "asis"}
invisible(lapply(names(cfas), function(cntr){
  print({kbl(cfas[[cntr]], caption = paste0("Scale reliability ", cntr)) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
})}))
```

We further examined measurement invariance across countries,
and found that metric invariance did not hold for these scales: `r paste0(colnames(meas_inv)[meas_inv[3, ] < .05], collapse = ", ")`.
This lack of measurement invariance must be taken into account when aggregating evidence across countries.

```{r tabinvar, results = "asis"}
meas_inv |>
  kbl(caption = "Measurement invariance tests for the difference between a configural and metrically invariant model.", digits = 3) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Planned analyses

```{r, code=xfun::read_utf8('../analysis.R')}

```

```{r tabresults, results = "asis"}
interpret_bf <- function(bf){
  sapply(bf, function(bf){
    if(bf < 1/10) return("rejected")
    if(bf < 1/3) return("rejected")
    if(bf >= 1/3 & bf <= 3) return("inconclusive")
    if(bf > 3 & bf <= 10) return("supported")
  return("supported")
  })
}
invisible(lapply(1:length(bayesfactors), function(h){
  print({
    kbl(est_for_hyp[[h]], caption = paste0("Parameters used to test hypothesis ", h, ". Evidence in favor of the hypothesis: BF = ", formatC(bayesfactors[h], digits = 3, format = "f"), " (", interpret_bf(bayesfactors[h]), ")."), digits = 3) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
})}))
```

According to these analyses, hypotheses `r paste0(which(bayesfactors > 3), collapse = ", ")` were supported.

These results are qualified by the relatively poor psychometric properties of some scales,
and the poor model fit of the CFA estimated in the three countries (see below).
Inspection of the modification indices suggested that adding cross-loadings might improve model fit,
but with the low number of indicators per factor this might compromise interpretability.
A likely explanation for the relatively poor model fit is the low explained variance in some items (see Table 3).

```{r tabfits, results = "asis"}
fit_17 <- t(sapply(res_list17, fitmeasures))
fit_17 <- fit_17[, c("npar", "chisq", "df", "cfi", "tli", "rmsea", "srmr")]
kbl(fit_17, caption = paste0("Fit of models used to test hypotheses 1-7."), digits = 3) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```