### 2021-12-16 manuscript/manuscript.Rmd
just omit the scale that prevents convergence. add meta-analysis

---
title: "Morality, politics, and cooperation"
output:
  bookdown::html_document2:
    keep_md: yes
date: '`r format(Sys.time(), "%d %B, %Y")`'
bibliography: references.bib
knit: worcs::cite_all
---

```{r setup, include=FALSE}
# We recommend that you prepare your raw data for analysis in 'prepare_data.R',
# and end that file with either open_data(yourdata), or closed_data(yourdata).
# Then, uncomment the line below to load the original or synthetic data
# (whichever is available), to allow anyone to reproduce your code:
library("worcs")
run_everything = FALSE
knitr::opts_chunk$set(echo = FALSE, results = "hide", message = FALSE, warning = FALSE)
options(knitr.kable.NA = '')
library(ggplot2)
library(tidySEM)
library(lavaan)
library(semTools)
library(kableExtra)
library(bookdown)
library(psych)
library(knitr)
dat <- load_data(to_envir = FALSE)[1:3]
source("../scales_list.R")
scales_list <- lapply(scales_list, function(i){
  tmp <- lapply(names(i), function(j){ paste0(j, "_", 1:length(i[[j]]))})
  names(tmp) <- names(i)
  tmp
  })
```

```{r cfas, eval = run_everything}
set.seed(1)
cfas <- lapply(names(scales_list), function(cnt){
  results <- t(sapply(scales_list[[cnt]], function(scal){
    out <- cfa(paste0("F =~ ", paste0(scal, collapse = "+")), std.lv = TRUE, data = dat[[cnt]])
    paran <- fa.parallel(dat[[cnt]][, scal])
    tab <- table_results(out, columns = NULL, digits=3)
    tab <- tab[-nrow(tab), ]
    loadns <- as.numeric(tab$est_std)[tab$op == "=~"]
    r2s <- lavInspect(out, "r2")
    omga <- semTools::reliability(out, what = "omega")[1,1]
    c(Subscale = gsub("_\\d$", "", scal[1]),
      Items = length(scal),
      n = out@Data@nobs[[1]],
      formatC(unlist(table_fit(out)[1, c("chisq", "cfi", "tli", "rmsea", "srmr")]), digits = 3, format = "f"), 
      formatC(c(min_load = min(loadns), max_load = max(loadns), min_r2 = min(r2s), max_r2 = max(r2s), omega = omga), digits = 3, format = "f"),
      Reliability = tidySEM:::interpret(omga),
      Factors = fa.parallel(dat[[cnt]][,scal]) $nfact
      )
  }))
})
names(cfas) <- names(scales_list)
saveRDS(cfas, "cfas.RData")
```
```{r}
cfas <- readRDS("cfas.RData")
```

## Necessary deviations from preregistration

When attempting to conduct the preregistered analyses,
the model did not converge in one of the datasets (NL).
We examined individual CFAs for the included scales to determine potential
sources of misspecification.
These analyses indicated that the scale `secs_eco` in that dataset
had no factors with Eigenvalues greater than would be expected by random chance (see column 'Factors', which is based on Horn's parallel analysis, 1965),
and had poor reliability (estimated using McDonald's Omega, which is calculated from the factor loadings and does not assume that all factor loadings are identical as Cronbach's alpha does).
After deleting this scale, the CFA model converged.
It was therefore deemed necessary to remove this scale.

```{r tabscale, results = "asis"}
invisible(lapply(names(cfas), function(cntr){
  print({kbl(cfas[[cntr]], caption = paste0("Scale reliability ", cntr)) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
})}))
```

<!-- To understand these issues better, we performed exploratory factor analysis on scales that showed indications of being non-unidimensional in at least one country. -->

```{r notunidim, results = "asis", include = FALSE}
allscales <- do.call(c, scales_list)
nfacs <- as.integer(do.call(c, lapply(cfas, function(i){i[, 'Factors']})))
check_these <- nfacs > 1
these_scales <- unique(do.call(c, lapply(cfas, function(i){i[, 'Subscale']}))[check_these])
check_these <- grep(paste0("^.{2}\\.(", paste0(these_scales, collapse = "|"), ")$"), names(allscales))
these_scales <- allscales[check_these]
nfacs <- nfacs[check_these]
nfacs[nfacs == 0] <- 2
out <- mapply(function(scal, nfac){
  fa(dat[[substring(scal, first = 1, last = 2)]][these_scales[[scal]]], nfac)
}, scal = names(these_scales), nfac = rep(2, length(these_scales)))

these_scales <- unique(do.call(c, lapply(cfas, function(i){i[, 'Subscale']}))[check_these])
for(scal in these_scales){
  print(kbl(do.call(cbind,
  lapply(grep(scal, names(out), value = TRUE), function(i){
    tmp <- data.frame(unclass(out[[i]][["loadings"]]))
    names(tmp) <- paste0(substring(i, 1,2), " Factor ", 1:ncol(tmp))
    rownames(tmp) <- paste0("Item ", 1:nrow(tmp))
    tmp
  })
), digits = 2, caption = paste0("Factor loadings for ", scal))|>
  kable_styling(bootstrap_options = c("striped", "hover")))
}
```

<!-- These factor analyses suggest that items 3-5 of sepa_soc loaded most consistently on the same factor. -->
<!-- Similarly, items 1-3 of sepa_eco loaded most consistently on the same factor. -->
<!-- For secs_soc, only items 4-5 loaded consistently high on the same factor. -->
<!-- For secs_eco, no items loaded consistently high on the same factor. -->
<!-- We removed items not consistently loading on one factor, -->
<!-- and dropped secs_eco entirely. -->
<!-- We applied these changes consistently across countries. -->

```{r, include=FALSE}
use_scales <- scales_list
use_scales$nl$secs_eco <- NULL
```

```{r, eval = run_everything}
set.seed(1)
cfas <- lapply(names(use_scales), function(cnt){
  results <- t(sapply(use_scales[[cnt]], function(scal){
    out <- cfa(paste0("F =~ ", paste0(scal, collapse = "+")), std.lv = TRUE, data = dat[[cnt]])
    paran <- fa.parallel(dat[[cnt]][, scal])
    tab <- table_results(out, columns = NULL, digits=3)
    tab <- tab[-nrow(tab), ]
    loadns <- as.numeric(tab$est_std)[tab$op == "=~"]
    r2s <- lavInspect(out, "r2")
    omga <- semTools::reliability(out, what = "omega")[1,1]
    c(Subscale = gsub("_\\d$", "", scal[1]),
      Items = length(scal),
      n = out@Data@nobs[[1]],
      formatC(unlist(table_fit(out)[1, c("chisq", "cfi", "tli", "rmsea", "srmr")]), digits = 3, format = "f"), 
      formatC(c(min_load = min(loadns), max_load = max(loadns), min_r2 = min(r2s), max_r2 = max(r2s), omega = omga), digits = 3, format = "f"),
      Reliability = tidySEM:::interpret(omga),
      Factors = invisible(fa.parallel(dat[[cnt]][,scal])$nfact)
      )
  }))
})
names(cfas) <- names(use_scales)

meas_inv <- sapply(use_scales$dk, function(scal){
  tmpdat <- rbind(tryCatch(cbind(dat$nl[, scal], Country = "NL"), error = function(e) NULL),
                  tryCatch(cbind(dat$dk[, scal], Country = "DK"), error = function(e) NULL),
                  tryCatch(cbind(dat$us[, scal], Country = "US"), error = function(e) NULL))
  conf <- cfa(paste0("F =~ ", paste0(scal, collapse = "+")), std.lv = TRUE, data = tmpdat, group = "Country")
  metr <- cfa(paste0("F =~ ", paste0(scal, collapse = "+")), std.lv = TRUE, data = tmpdat, group = "Country", group.equal = "loadings")
  tst <- anova(conf, metr)
  unlist(tst[2, c("Chisq diff", "Df diff", "Pr(>Chisq)")])
})
saveRDS(cfas, "cfas2.RData")
saveRDS(meas_inv, "meas_inv.RData")
```

```{r tabscaleuse, results = "asis"}
cfas <- readRDS("cfas2.RData")
meas_inv <- readRDS("meas_inv.RData")
invisible(lapply(names(cfas), function(cntr){
  print({kbl(cfas[[cntr]], caption = paste0("Scale reliability ", cntr)) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
})}))
```

We further examined measurement invariance across countries,
and found that metric invariance did not hold for these scales: `r paste0(colnames(meas_inv)[meas_inv[3, ] < .05], collapse = ", ")`.
This lack of measurement invariance must be taken into account when aggregating evidence across countries.

```{r tabinvar, results = "asis"}
meas_inv |>
  kbl(caption = "Measurement invariance tests for the difference between a configural and metrically invariant model.", digits = 3) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Planned analyses

```{r, eval = run_everything}
library(worcs)
library(lavaan)
library(BFpack)
library(ggplot2)
library(tidySEM)
library(metafor)
set.seed(22)
sample_sizes <- sapply(dat, nrow)
hypotheses <- list("(fam_secs_soc_us, fam_secs_eco_us, fam_sepa_soc_dk, fam_sepa_eco_dk, fam_secs_soc_nl, fam_secs_eco_nl, fam_sepa_soc_nl, fam_sepa_eco_nl) > .1",
                   "(grp_secs_soc_us, grp_secs_eco_us, grp_sepa_soc_dk, grp_sepa_eco_dk, grp_secs_soc_nl, grp_secs_eco_nl, grp_sepa_soc_nl, grp_sepa_eco_nl) > .1",
                   "(rec_secs_soc_us, rec_secs_eco_us, rec_sepa_soc_dk, rec_sepa_eco_dk, rec_secs_soc_nl, rec_secs_eco_nl, rec_sepa_soc_nl, rec_sepa_eco_nl) > .1",
                   "(her_secs_soc_us, her_secs_eco_us, her_sepa_soc_dk, her_sepa_eco_dk, her_secs_soc_nl, her_secs_eco_nl, her_sepa_soc_nl, her_sepa_eco_nl) > .1",
                   "(def_secs_soc_us, def_secs_eco_us, def_sepa_soc_dk, def_sepa_eco_dk, def_secs_soc_nl, def_secs_eco_nl, def_sepa_soc_nl, def_sepa_eco_nl) > .1",
                   "(fai_secs_soc_us, fai_secs_eco_us, fai_sepa_soc_dk, fai_sepa_eco_dk, fai_secs_soc_nl, fai_secs_eco_nl, fai_sepa_soc_nl, fai_sepa_eco_nl) < -.1",
                   "(pro_secs_soc_us, pro_secs_eco_us, pro_sepa_soc_dk, pro_sepa_eco_dk, pro_secs_soc_nl, pro_secs_eco_nl, pro_sepa_soc_nl, pro_sepa_eco_nl) > .1"
                   )
hypotheses[[8]] <- gsub("[<>]", "=", gsub("\\b(fam|grp)_((sepa|secs)_(soc|eco))_", "\\2ON\\1_", gsub("\\) > 0 & \\(", ", ", paste0(gsub("-?\\.1", "0", hypotheses[[1]]), " & ", gsub(".1", "0", hypotheses[[2]], fixed = TRUE)))))
hypotheses[[9]] <- gsub("fam", "rec", gsub("grp", "fai", hypotheses[[8]], fixed = TRUE), fixed = TRUE)

mods_cor <- lapply(names(dat), function(country){
  cntdat <- dat[[country]][unlist(use_scales[[country]])]
  mod <- tidy_sem(cntdat)
  measurement(mod)
})
names(mods_cor) <- names(dat)
mods_8 <- lapply(mods_cor, function(x){
  add_paths(x,   paste0((t(outer(paste0(unique(grep("^se", x$dictionary$scale, value = TRUE)), " ~ "),
                                 unique(grep("^(fam|grp)", x$dictionary$scale, value = TRUE)), paste0))), collapse = "\n"))
})
mods_9 <- lapply(mods_cor, function(x){
  add_paths(x,   paste0((t(outer(paste0(unique(grep("^se", x$dictionary$scale, value = TRUE)), " ~ "),
                                 unique(grep("^(rec|fai)", x$dictionary$scale, value = TRUE)), paste0))), collapse = "\n"))
})

removeothercountries <- function(hyp, cnt){
  gsub(",\\s+\\)", "\\)", gsub(paste0("\\b[a-zA-Z_]+(?<!", cnt, ")\\b,?"), "", hyp, perl = TRUE))
}

removeparameters <- function(hyp, est){
  params_in_hyp <- bain:::params_in_hyp(hyp)
  est <- names(est)
  if(any(!params_in_hyp %in% est)){
    for(param in params_in_hyp[!params_in_hyp %in% est]){
      hyp <- gsub(paste0("\\b", param, "\\b,?"), "", hyp)
    }
    hyp <- gsub(",\\s+\\)", "\\)", hyp)
    hyp <- gsub("\\(\\s+", "\\(", hyp)
  }
  hyp <- gsub("\\(([a-zA-Z_]+)\\)", "\\1", hyp)
  if(length(bain:::params_in_hyp(hyp)) == 0) return(NA)
  hyp
}

res_list <- lapply(mods_cor, function(x){estimate_lavaan(x)})
res_list17 <- res_list
res <- lapply(names(res_list), function(country){
  thisfit = res_list[[country]]
  tab <- bain:::lav_get_estimates(thisfit, standardize = TRUE, retain_which = "~~")
  keep_these <- which(!grepl("~~(secs|sepa)_(soc|eco)$", names(tab$estimate)) & grepl("(secs|sepa)_(soc|eco)~~", names(tab$estimate)))
  estimates <- tab$estimate[keep_these]
  Sigma <- tab$Sigma[[1]][keep_these, keep_these]
  names(estimates) <- paste0(gsub("~~", "_", names(estimates), fixed = TRUE), "_", country)
  names(estimates) <- gsub("^(sepa|secs)_(.+?)_(.+?)_(.+)$", "\\3_\\1_\\2_\\4", names(estimates))
  colnames(Sigma) <- rownames(Sigma) <- names(estimates)
  list(est = estimates, sig = Sigma)
})
names(res) <- names(res_list)
res_h17 <- res
# Get individual BFs for each sample
bf_individual <- sapply(1:length(res_list), function(i){
  hyp <- paste2(unlist(
    lapply(
      lapply(hypotheses[1:7], removeothercountries, cnt = names(res_list)[i]),
      removeparameters, est = res[[i]]$est)
    ), collapse = "; ")
  tmp <- BF(x = res[[i]]$est,
            hypothesis = hyp,
            Sigma = res[[i]]$sig,
            n = sample_sizes[i])
  tmp$BFtu_confirmatory[1:7]
})
out_cors <- apply(bf_individual, 1, prod)

out_8 <- 
  {
    # fit model
    res_list <- lapply(mods_8, function(x){estimate_lavaan(x)})
    res_list8 <- res_list
    res <- lapply(names(res_list), function(country){
      thisfit = res_list[[country]]
      tab <- bain:::lav_get_estimates(thisfit, standardize = TRUE, retain_which = "~")
      estimates <- tab$estimate
      Sigma <- tab$Sigma[[1]]
      names(estimates) <- paste0(gsub("~", "ON", names(estimates), fixed = TRUE), "_", country)
      colnames(Sigma) <- rownames(Sigma) <- names(estimates)
      list(est = estimates, sig = Sigma)
    })
    res8 <- res
    bf_individual8 <- sapply(1:length(res_list), function(i){
      hyp <- removeparameters(removeothercountries(hypotheses[8], cnt = names(res_list)[i]), est = res[[i]]$est)
      tmp <- BF(x = res[[i]]$est,
                hypothesis = hyp,
                Sigma = res[[i]]$sig,
                n = sample_sizes[i])
      tmp$BFtu_confirmatory[1]
    })
    prod(bf_individual8)
  }
out_9 <- 
  {
    # fit model
    res_list <- lapply(mods_9, function(x){estimate_lavaan(x)})
    res_list9 <- res_list
    res <- lapply(names(res_list), function(country){
      thisfit = res_list[[country]]
      tab <- bain:::lav_get_estimates(thisfit, standardize = TRUE, retain_which = "~")
      estimates <- tab$estimate
      Sigma <- tab$Sigma[[1]]
      names(estimates) <- paste0(gsub("~", "ON", names(estimates), fixed = TRUE), "_", country)
      colnames(Sigma) <- rownames(Sigma) <- names(estimates)
      list(est = estimates, sig = Sigma)
    })
    res9 <- res
    bf_individual9 <- sapply(1:length(res_list), function(i){
      hyp <- removeparameters(removeothercountries(hypotheses[9], cnt = names(res_list)[i]), est = res[[i]]$est)
      tmp <- BF(x = res[[i]]$est,
                hypothesis = hyp,
                Sigma = res[[i]]$sig,
                n = sample_sizes[i])
      tmp$BFtu_confirmatory[1]
    })
    prod(bf_individual9)
  }
bayesfactors <- c(out_cors, 1/out_8, 1/out_9)
est_for_hyp <- lapply(hypotheses, function(h){
  pars = bain:::params_in_hyp(h)
  tmp <- unlist(lapply(c(res_h17, res8, res9), function(i){
    i$est[which(names(i$est) %in% pars)]
  }))
  nams <- names(tmp)
  tmp <- data.frame(Value = tmp, Country = gsub("^.+(.{2})$", "\\1", nams), Parameter = gsub("^(.{2}\\.)?(.+?)_.{2}$", "\\2", nams))
  tmp <- reshape(tmp, direction = "wide", idvar = "Country", timevar =  "Parameter")
  rownames(tmp) <- NULL
  names(tmp) <- gsub("Value\\.", "", names(tmp))
  tmp
})

estimates <- lapply(names(res_list17), function(c){
    x <- res_list17[[c]]
    tmp <- standardizedSolution(x)
    tmp <- tmp[tmp$op == "~~" & grepl("(soc|eco)", tmp$lhs) & !grepl("(soc|eco)", tmp$rhs), c("rhs", "lhs", "est.std", "se")]
    names(tmp)[3] <- "ri"
    tmp$social <- as.integer(grepl("_soc", tmp$lhs, fixed = TRUE))
    tmp$eco <- as.integer(grepl("_eco", tmp$lhs, fixed = TRUE))
    tmp$country <- c
    tmp$ni <- unlist(x@Data@nobs)
    tmp
})
estimates <- do.call(rbind, estimates)

meta17 <- lapply(unique(estimates$rhs), function(x){
  tmp <- estimates[estimates$rhs==x,]
  tmp <- escalc(ri = tmp$ri, ni = tmp$ni, measure = "ZCOR", data = tmp)
  # res <- rma(tmp$yi, vi = tmp$vi, mods = tmp[, c("social", "eco")], intercept = FALSE)
  # if(length(unique(tmp$country))>2){
  #   res <- robust(res, cluster = tmp$country)  
  # }
  tmp$id_study <- tmp$country
  tmp$id_es <- 1:nrow(tmp)
  res <- rma.mv(tmp$yi, tmp$vi, random = list(~ 1 | id_study, ~ 1 | id_es), data=tmp) 
  confints <- confint(res)
  tabres <- data.frame(Parameter = c(rownames(res$b), "Tau2b", "Tau2w"),
                       Estimate = c(fisherz2r(res$b), res$sigma2),
                       ci.lb = c(fisherz2r(res$ci.lb), confints[[1]]$random[1,2], confints[[2]]$random[1,2]),
                       ci.ub = c(fisherz2r(res$ci.ub), confints[[1]]$random[1,3], confints[[2]]$random[1,3]),
                       p = c(res$pval, NA, NA),
                       sig = c(c("", "*")[(res$pval < .05)+1], c("", "*")[(c(confints[[1]]$random[1,2], confints[[2]]$random[1,2])> 1e-4)+1]),
                       Model = "Overall")
  # By conservatism dimension
  res <- rma.mv(tmp$yi, tmp$vi, random = list(~ 1 | id_study, ~ 1 | id_es), data=tmp, intercept = FALSE, mods = tmp[, c("social", "eco")]) 
  confints <- confint(res)
  tabres2 <- data.frame(Parameter = c(rownames(res$b), "Tau2b", "Tau2w"),
                       Estimate = c(fisherz2r(res$b), res$sigma2),
                       ci.lb = c(fisherz2r(res$ci.lb), confints[[1]]$random[1,2], confints[[2]]$random[1,2]),
                       ci.ub = c(fisherz2r(res$ci.ub), confints[[1]]$random[1,3], confints[[2]]$random[1,3]),
                       p = c(res$pval, NA, NA),
                       sig = c(c("", "*")[(res$pval < .05)+1], c("", "*")[(c(confints[[1]]$random[1,2], confints[[2]]$random[1,2])> 1e-4)+1]),
                       Model = "Multigroup")
  tabres <- rbind(tabres, tabres2)
  tabres$CI <- conf_int(lb = tabres$ci.lb, ub = tabres$ci.ub)
  tabres[c("ci.lb", "ci.ub")] <- NULL
  tabres$Domain <- x
  tabres
})
meta17 <- do.call(rbind, meta17)

# Hypothesis 8 ------------------------------------------------------------
source("mod89.R")
res8 <- lapply(names(mod8), function(n){
  sem(mod8[[n]], dat[[n]], auto.fix.first = FALSE, auto.var = TRUE, auto.cov.lv.x = TRUE, std.lv = TRUE, auto.cov.y = TRUE)
})
names(res8) <- names(mod8)

estimates <- lapply(names(res8), function(c){
  x <- res8[[c]]
  tmp <- parameterestimates(x)
  tmp <- tmp[!tmp$label == "", c("rhs", "lhs", "est", "se")]
  names(tmp)[3] <- "bi"
  tmp$social <- as.integer(grepl("_soc", tmp$lhs, fixed = TRUE))
  tmp$eco <- as.integer(grepl("_eco", tmp$lhs, fixed = TRUE))
  tmp$country <- c
  tmp$ni <- unlist(x@Data@nobs)
  tmp
})
estimates <- do.call(rbind, estimates)

meta8 <- lapply(c("fam", "grp"), function(x){
  tmp <- estimates[estimates$rhs==x,]
  tmp$id_study <- tmp$country
  tmp$id_es <- 1:nrow(tmp)
  if(nrow(tmp) == 1) return(NULL)
  res <- tryCatch({rma.mv(tmp$bi, tmp$se^2, random = list(~ 1 | id_study, ~ 1 | id_es), data=tmp) }, error = function(e){rma(tmp$bi, tmp$se^2, data=tmp)})
  confints <- confint(res)
  tabres <- data.frame(Parameter = c(rownames(res$b), "Tau2b", "Tau2w"),
                       Estimate = c(fisherz2r(res$b), res$sigma2),
                       ci.lb = c(fisherz2r(res$ci.lb), confints[[1]]$random[1,2], confints[[2]]$random[1,2]),
                       ci.ub = c(fisherz2r(res$ci.ub), confints[[1]]$random[1,3], confints[[2]]$random[1,3]),
                       p = c(res$pval, NA, NA),
                       sig = c(c("", "*")[(res$pval < .05)+1], c("", "*")[(c(confints[[1]]$random[1,2], confints[[2]]$random[1,2])> 1e-4)+1]),
                       Model = "Overall")
  # By conservatism dimension
  res <- try(rma.mv(tmp$bi, tmp$se^2, random = list(~ 1 | id_study, ~ 1 | id_es), data=tmp, intercept = FALSE, mods = tmp[, c("social", "eco")]))
  if(inherits(res, "try-error")) return(tabres)
  confints <- confint(res)
  tabres2 <- data.frame(Parameter = c(rownames(res$b), "Tau2b", "Tau2w"),
                        Estimate = c(fisherz2r(res$b), res$sigma2),
                        ci.lb = c(fisherz2r(res$ci.lb), confints[[1]]$random[1,2], confints[[2]]$random[1,2]),
                        ci.ub = c(fisherz2r(res$ci.ub), confints[[1]]$random[1,3], confints[[2]]$random[1,3]),
                        p = c(res$pval, NA, NA),
                        sig = c(c("", "*")[(res$pval < .05)+1], c("", "*")[(c(confints[[1]]$random[1,2], confints[[2]]$random[1,2])> 1e-4)+1]),
                        Model = "Multigroup")
  if(tabres$Estimate[1] == tabres2$Estimate[1]) return(tabres)
  tabres <- rbind(tabres, tabres2)
  tabres$CI <- conf_int(lb = tabres$ci.lb, ub = tabres$ci.ub)
  tabres[c("ci.lb", "ci.ub")] <- NULL
  tabres$Domain <- x
  tabres
})
meta8 <- do.call(rbind, meta8)

# Hypothesis 9 ------------------------------------------------------------

res9 <- lapply(names(mod8), function(n){
  mod <- mod8[[n]]
  mod <- gsub("secs_soc_fam", "secs_soc_rec", mod, fixed = T)
  mod <- gsub("secs_soc_grp", "secs_soc_fai", mod, fixed = T)
  mod <- gsub("* fam", "* rec", mod, fixed = T)
  mod <- gsub("* grp", "* fai", mod, fixed = T)
  sem(mod, dat[[n]], auto.fix.first = FALSE, auto.var = TRUE, auto.cov.lv.x = TRUE, std.lv = TRUE, auto.cov.y = TRUE)
})
names(res9) <- names(mod8)

estimates <- lapply(names(res9), function(c){
  x <- res9[[c]]
  tmp <- parameterestimates(x)
  tmp <- tmp[!tmp$label == "", c("rhs", "lhs", "est", "se")]
  names(tmp)[3] <- "bi"
  tmp$social <- as.integer(grepl("_soc", tmp$lhs, fixed = TRUE))
  tmp$eco <- as.integer(grepl("_eco", tmp$lhs, fixed = TRUE))
  tmp$country <- c
  tmp$ni <- unlist(x@Data@nobs)
  tmp
})
estimates <- do.call(rbind, estimates)

meta9 <- lapply(c("rec", "fai"), function(x){
  tmp <- estimates[estimates$rhs==x,]
  tmp$id_study <- tmp$country
  tmp$id_es <- 1:nrow(tmp)
  if(nrow(tmp) == 1) return(NULL)
  res <- tryCatch({rma.mv(tmp$bi, tmp$se^2, random = list(~ 1 | id_study, ~ 1 | id_es), data=tmp) }, error = function(e){rma(tmp$bi, tmp$se^2, data=tmp)})
  confints <- confint(res)
  tabres <- data.frame(Parameter = c(rownames(res$b), "Tau2b", "Tau2w"),
                       Estimate = c(fisherz2r(res$b), res$sigma2),
                       ci.lb = c(fisherz2r(res$ci.lb), confints[[1]]$random[1,2], confints[[2]]$random[1,2]),
                       ci.ub = c(fisherz2r(res$ci.ub), confints[[1]]$random[1,3], confints[[2]]$random[1,3]),
                       p = c(res$pval, NA, NA),
                       sig = c(c("", "*")[(res$pval < .05)+1], c("", "*")[(c(confints[[1]]$random[1,2], confints[[2]]$random[1,2])> 1e-4)+1]),
                       Model = "Overall")
  # By conservatism dimension
  res <- try(rma.mv(tmp$bi, tmp$se^2, random = list(~ 1 | id_study, ~ 1 | id_es), data=tmp, intercept = FALSE, mods = tmp[, c("social", "eco")]))
  if(inherits(res, "try-error")) return(tabres)
  confints <- confint(res)
  tabres2 <- data.frame(Parameter = c(rownames(res$b), "Tau2b", "Tau2w"),
                        Estimate = c(fisherz2r(res$b), res$sigma2),
                        ci.lb = c(fisherz2r(res$ci.lb), confints[[1]]$random[1,2], confints[[2]]$random[1,2]),
                        ci.ub = c(fisherz2r(res$ci.ub), confints[[1]]$random[1,3], confints[[2]]$random[1,3]),
                        p = c(res$pval, NA, NA),
                        sig = c(c("", "*")[(res$pval < .05)+1], c("", "*")[(c(confints[[1]]$random[1,2], confints[[2]]$random[1,2])> 1e-4)+1]),
                        Model = "Multigroup")
  if(tabres$Estimate[1] == tabres2$Estimate[1]) return(tabres)
  tabres <- rbind(tabres, tabres2)
  tabres$CI <- conf_int(lb = tabres$ci.lb, ub = tabres$ci.ub)
  tabres[c("ci.lb", "ci.ub")] <- NULL
  tabres$Domain <- x
  tabres
})
meta9 <- do.call(rbind, meta9)

saveRDS(bayesfactors, "bayesfactors.RData")
saveRDS(est_for_hyp, "est_for_hyp.RData")
saveRDS(res_list17, "res_list17.RData")
saveRDS(meta17, "meta17.RData")
saveRDS(meta8, "meta8.RData")
saveRDS(meta9, "meta9.RData")

```

First, we provide an overview of the latent variable correlations for each country.
These are the model parameters evaluated using the informative hypothesis tests below:

```{r templatecortabs, include=FALSE}
res_list17 <- readRDS("res_list17.RData")
out = NULL
for (i in 1:length(res_list17)) {
  out = c(out, knit_expand('template_tabcor.rmd'))
}
```

`r paste(knit(text = out), collapse = '\n')`

Next, we examine evidence for the preregistered hypotheses:

```{r tabresults, results = "asis"}
bayesfactors <- readRDS("bayesfactors.RData")
est_for_hyp <- readRDS("est_for_hyp.RData")
interpret_bf <- function(bf){
  sapply(bf, function(bf){
    if(bf < 1/10) return("rejected")
    if(bf < 1/3) return("rejected")
    if(bf >= 1/3 & bf <= 3) return("inconclusive")
    if(bf > 3 & bf <= 10) return("supported")
  return("supported")
  })
}
invisible(lapply(1:length(bayesfactors), function(h){
  print({
    kbl(est_for_hyp[[h]], caption = paste0("Parameters used to test hypothesis ", h, ". Evidence in favor of the hypothesis: BF = ", formatC(bayesfactors[h], digits = 3, format = "f"), " (", interpret_bf(bayesfactors[h]), ")."), digits = 3) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
})}))
```

According to these analyses, hypotheses `r paste0(which(bayesfactors > 3), collapse = ", ")` were supported.

These results are qualified by the relatively poor psychometric properties of some scales,
the lack of measurement invariance for some scales,
and the poor model fit of the CFA estimated in the three countries (see below).
Inspection of the modification indices suggested that adding cross-loadings might improve model fit,
but with the low number of indicators per factor this might compromise interpretability.
A likely explanation for the relatively poor model fit is the low explained variance in some items (see Table 3).

```{r tabfits, results = "asis"}
fit_17 <- t(sapply(res_list17, fitmeasures))
fit_17 <- fit_17[, c("npar", "chisq", "df", "cfi", "tli", "rmsea", "srmr")]
kbl(fit_17, caption = paste0("Fit of models used to test hypotheses 1-7."), digits = 3) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Exploratory analyses

The preceding analyses were conceptually replicated in a fourth sample.
Note that this sample used moral relevance scales instead of moral judgment scales.

```{r, eval = run_everything}
library(worcs)
library(lavaan)
library(BFpack)
library(ggplot2)
library(tidySEM)
tris <- load_data(to_envir = FALSE)[["tris"]]

# Descriptives ------------------------------------------------------------
scales_tris <-
  lapply(unique(gsub("_\\d+$", "", names(tris))), function(i)
    grep(i, names(tris), value = T))
names(scales_tris) <- unique(gsub("_\\d+$", "", names(tris)))

cfa_tris <- t(sapply(names(scales_tris), function(scal) {
  scal <- paste0(scal, "_", 1:length(scales_tris[[scal]]))
  out <-
    cfa(paste0("F =~ ", paste0(scal, collapse = "+")), std.lv = TRUE, data = tris)
  paran <- fa.parallel(tris[, scal])
  tab <- table_results(out, columns = NULL, digits = 3)
  tab <- tab[-nrow(tab),]
  loadns <- as.numeric(tab$est_std)[tab$op == "=~"]
  r2s <- lavInspect(out, "r2")
  omga <- semTools::reliability(out, what = "omega")[1, 1]
  c(
    Subscale = gsub("_\\d$", "", scal[1]),
    Items = length(scal),
    n = out@Data@nobs[[1]],
    formatC(unlist(table_fit(out)[1, c("chisq", "cfi", "tli", "rmsea", "srmr")]), digits = 3, format = "f"),
    formatC(
      c(
        min_load = min(loadns),
        max_load = max(loadns),
        min_r2 = min(r2s),
        max_r2 = max(r2s),
        omega = omga
      ),
      digits = 3,
      format = "f"
    ),
    Reliability = tidySEM:::interpret(omga),
    Factors = paran$nfact
  )
}))

sample_sizes <- nrow(tris)
hypotheses <-
  list(
    "(fam_secs_soc_us, fam_secs_eco_us, fam_sepa_soc_dk, fam_sepa_eco_dk, fam_secs_soc_nl, fam_secs_eco_nl, fam_sepa_soc_nl, fam_sepa_eco_nl) > .1",
    "(grp_secs_soc_us, grp_secs_eco_us, grp_sepa_soc_dk, grp_sepa_eco_dk, grp_secs_soc_nl, grp_secs_eco_nl, grp_sepa_soc_nl, grp_sepa_eco_nl) > .1",
    "(rec_secs_soc_us, rec_secs_eco_us, rec_sepa_soc_dk, rec_sepa_eco_dk, rec_secs_soc_nl, rec_secs_eco_nl, rec_sepa_soc_nl, rec_sepa_eco_nl) > .1",
    "(her_secs_soc_us, her_secs_eco_us, her_sepa_soc_dk, her_sepa_eco_dk, her_secs_soc_nl, her_secs_eco_nl, her_sepa_soc_nl, her_sepa_eco_nl) > .1",
    "(def_secs_soc_us, def_secs_eco_us, def_sepa_soc_dk, def_sepa_eco_dk, def_secs_soc_nl, def_secs_eco_nl, def_sepa_soc_nl, def_sepa_eco_nl) > .1",
    "(fai_secs_soc_us, fai_secs_eco_us, fai_sepa_soc_dk, fai_sepa_eco_dk, fai_secs_soc_nl, fai_secs_eco_nl, fai_sepa_soc_nl, fai_sepa_eco_nl) < -.1",
    "(pro_secs_soc_us, pro_secs_eco_us, pro_sepa_soc_dk, pro_sepa_eco_dk, pro_secs_soc_nl, pro_secs_eco_nl, pro_sepa_soc_nl, pro_sepa_eco_nl) > .1"
  )
hypotheses[[8]] <-
  gsub("[<>]",
       "=",
       gsub(
         "\\b(fam|grp)_((sepa|secs)_(soc|eco))_",
         "\\2ON\\1_",
         gsub("\\) > 0 & \\(", ", ", paste0(
           gsub("-?\\.1", "0", hypotheses[[1]]),
           " & ",
           gsub(".1", "0", hypotheses[[2]], fixed = TRUE)
         ))
       ))
hypotheses[[9]] <-
  gsub("fam",
       "rec",
       gsub("grp", "fai", hypotheses[[8]], fixed = TRUE),
       fixed = TRUE)

mod <- tidy_sem(tris)
mod_cor <- measurement(mod)

mod_8 <-
  add_paths(mod_cor, paste0((t(
    outer(paste0(unique(
      grep("^se", mod_cor$dictionary$scale, value = TRUE)
    ), " ~ "),
    unique(
      grep("^(fam|grp)", mod_cor$dictionary$scale, value = TRUE)
    ), paste0)
  )), collapse = "\n"))

mod_9 <-
  add_paths(mod_cor,   paste0((t(
    outer(paste0(unique(
      grep("^se", mod_cor$dictionary$scale, value = TRUE)
    ), " ~ "),
    unique(
      grep("^(rec|fai)", mod_cor$dictionary$scale, value = TRUE)
    ), paste0)
  )), collapse = "\n"))

removeothercountries <- function(hyp, cnt) {
  gsub(",\\s+\\)", "\\)", gsub(paste0("\\b[a-zA-Z_]+(?<!", cnt, ")\\b,?"), "", hyp, perl = TRUE))
}

removeparameters <- function(hyp, est) {
  params_in_hyp <- bain:::params_in_hyp(hyp)
  est <- names(est)
  if (any(!params_in_hyp %in% est)) {
    for (param in params_in_hyp[!params_in_hyp %in% est]) {
      hyp <- gsub(paste0("\\b", param, "\\b,?"), "", hyp)
    }
    hyp <- gsub(",\\s+\\)", "\\)", hyp)
    hyp <- gsub("\\(\\s+", "\\(", hyp)
  }
  hyp <- gsub("\\(([a-zA-Z_]+)\\)", "\\1", hyp)
  if (length(bain:::params_in_hyp(hyp)) == 0)
    return(NA)
  hyp
}

res17 <- estimate_lavaan(mod_cor)
res_tris <- res17
tab <-
  bain:::lav_get_estimates(res17, standardize = TRUE, retain_which = "~~")
keep_these <-
  which(
    !grepl("~~(secs|sepa)_(soc|eco)$", names(tab$estimate)) &
      grepl("(secs|sepa)_(soc|eco)~~", names(tab$estimate))
  )
estimates <- tab$estimate[keep_these]
Sigma <- tab$Sigma[[1]][keep_these, keep_these]
names(estimates) <-
  gsub("~~", "_", names(estimates), fixed = TRUE)
names(estimates) <-
  gsub("^(sepa|secs)_(.+?)_(.+?)$", "\\3_\\1_\\2", names(estimates))
colnames(Sigma) <- rownames(Sigma) <- names(estimates)
res <- list(est = estimates, sig = Sigma)
res17 <- res
# Get BF
hyp <- paste2(unlist(lapply(
  gsub("_us", "", lapply(hypotheses[1:7], removeothercountries, cnt = "us")),
  removeparameters, est = res$est
)), collapse = "; ")
tmp <- BF(
  x = res$est,
  hypothesis = hyp,
  Sigma = res$sig,
  n = sample_sizes
)

out_cors <- tmp$BFtu_confirmatory[1:7]

out_8 <-
  {
    # fit model
    res <- estimate_lavaan(mod_8)
    res <- {
      thisfit = res
      tab <-
        bain:::lav_get_estimates(thisfit, standardize = TRUE, retain_which = "~")
      estimates <- tab$estimate
      Sigma <- tab$Sigma[[1]]
      names(estimates) <-
        gsub("~", "ON", names(estimates), fixed = TRUE)
      colnames(Sigma) <- rownames(Sigma) <- names(estimates)
      list(est = estimates, sig = Sigma)
    }
    res_tris8 <- res

    hyp <- paste2(removeparameters(gsub(
      "_us",
      "",
      lapply(hypotheses[8], removeothercountries, cnt = "us")
    ),
    est = res$est))
    tmp <- BF(
      x = res$est,
      hypothesis = hyp,
      Sigma = res$sig,
      n = sample_sizes
    )
    tmp$BFtu_confirmatory[1]

  }
out_9 <-
  {
    # fit model
    res <- estimate_lavaan(mod_9)
    res <- {
      thisfit = res
      tab <-
        bain:::lav_get_estimates(thisfit, standardize = TRUE, retain_which = "~")
      estimates <- tab$estimate
      Sigma <- tab$Sigma[[1]]
      names(estimates) <-
        gsub("~", "ON", names(estimates), fixed = TRUE)
      colnames(Sigma) <- rownames(Sigma) <- names(estimates)
      list(est = estimates, sig = Sigma)
    }
    res_tris9 <- res

    hyp <- paste2(removeparameters(gsub(
      "_us",
      "",
      lapply(hypotheses[9], removeothercountries, cnt = "us")
    ),
    est = res$est))
    tmp <- BF(
      x = res$est,
      hypothesis = hyp,
      Sigma = res$sig,
      n = sample_sizes
    )
    tmp$BFtu_confirmatory[1]

  }

bayesfactors_tris <- c(out_cors, 1 / out_8, 1 / out_9)
est_for_hyp_tris <- lapply(hypotheses, function(h) {
  pars = gsub("_us", "", bain:::params_in_hyp(h))
  tmp <- unlist(lapply(list(res17, res_tris8, res_tris9), function(i) {
    tryCatch(i$est[which(names(i$est) %in% pars)], error = function(e){ NA })
  }))
  tmp
})

saveRDS(bayesfactors_tris, "bayesfactors_tris.RData")
saveRDS(est_for_hyp_tris, "est_for_hyp_tris.RData")
saveRDS(res_tris, "res_tris.RData")
```

```{r}
bayesfactors_tris <- readRDS("bayesfactors_tris.RData")
est_for_hyp_tris <- readRDS("est_for_hyp_tris.RData")
res_tris <- readRDS("res_tris.RData")
```

This replication offered no support for any hypothesis except H`r which(bayesfactors_tris > 3)`:

```{r, results = "asis"}
trisfit <- unlist(table_fit(res_tris)[1, c("chisq", "cfi", "tli", "rmsea", "srmr")])
for(i in 1:length(bayesfactors_tris)){
  cat(
    "* $BF_{h", i, "} = ", formatC(bayesfactors_tris[i], digits = 3, format = "f"),
    "$, based on parameters ", 
    paste0(names(est_for_hyp_tris[[i]]), " (", formatC(est_for_hyp_tris[[i]], digits = 3, format = "f"), ")", collapse = ", "), ".  \r\n",
    sep = ""
  )
}
```

The fit of this model was also poor, though somewhat better than for the confirmatory analyses; `r paste0(names(trisfit), " = ", formatC(trisfit, digits = 3, format = "f"), collapse = ", ")`.

# Meta-analyses

Whereas the preregistered hypotheses test the hypothesis that *all* correlations between *all* moral domains and *all* conservatism scales will be > .1 in *all* countries,
we can also examine the *average* correlation between moral domains and conservatism scales across countries.

For this purpose, we conducted three-level meta-analyses.
We estimated two separate models for each moral domain:
One with an average correlation (intercept) across all conservatism scales (Model: Overall),
and one multi-group analysis with separate intercepts for social and economic conservatism scales (Model: Multigroup).
To account for the dependent data, we used a three-level meta-analysis.
Note that each model provides an estimate of variance between effect sizes within each country (tau$^2$w) and variance between the countries (tau$^2$b).

```{r, results = "asis"}
meta17 <- readRDS("meta17.RData")
kbl(meta17, caption = "Average correlations according to three-level meta-analysis", digits = 3) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

For Hypotheses 8 and 9, it is instructive to examine the average partial regression coefficients for the two moral domains mentioned in the hypotheses:

```{r meta8, results = "asis"}
meta8 <- readRDS("meta8.RData")
kbl(meta8, caption = "Average partial regression coefficients according to three-level meta-analysis", digits = 3) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r meta9, results = "asis"}
meta9 <- readRDS("meta9.RData")
kbl(meta9, caption = "Average partial regression coefficients according to three-level meta-analysis", digits = 3) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```